Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex

### 2019-04-01 Manuscript.Rmd
Removed density files due to miscalculation (fixed in corrCounts), plus revisions to manuscript

Our fMRI analyses used resting-state fMRI data from the Human Connectome Project [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state runs (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) each comprised 1200 time points(14 min 33 s) for a total of 4800 time points. Two runs were acquired on each day, with the phase encoding direction set to left-right for one run and right-left for the other. Only subjects with both left-right and right-left phase encoding were included (i.e. subjects with four rsfMRI sessions). In addition, only datasets with either low motion levels (under 1.5mm) or under 0.5 mm mean framewise displacement (FD) were used. See [@VanEssen2012] for more details about the data acquisition protocol.

@@

```{r Density calculation, out.width="60%", fig.align = "center"}
# Load density data
setwd('./Densities/')
temp <- list.files()
Densities <- do.call(rbind, lapply(temp, read.csv))

# Get the mean density for all data, daily scans, and sessions
mdAll <- mean(Densities$Density[grep("Overall", Densities$Type)])
sddAll <- sd(Densities$Density[grep("Overall", Densities$Type)])
mdDay <- mean(Densities$Density[grep("Day", Densities$Type)])
mdSess <- mean(Densities$Density[grep("Session", Densities$Type)])
```

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). **TODO: recompute corrCounts so it also stores the # of pos/neg correlations turned to 0, as well as the number of TRs per subject (add to fMRI details)**. Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. The proportion of remaining edges (i.e. network density) was high for all individuals(mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). We generated and analyzed network weight matrices at four levels: (1) for each run separately (1200 TRs each); (2) for the concatenated time series for the two runs on each day (2400 TRs total); (3) for each subject's full dataset (4800 TRs total); and (4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

In order to evaluate the convergent validity of the resulting partitionings across community-detection methods, we also estimated these communities using the more traditional approach of modularity maximization, based on the algorithm from Clauset et al. [-@Clauset2004]. **(TODO: rephrase) This method maximizes a quality function that compares the strength of the connection between any two vertices against a null model of their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the clustering that maximizes the network's modularity.** Unlike SP, modularity can fractionate a network into more than two communities. We assumed that if the partitionings from the bounded and unbounded community detection methods showed high levels of agreement, we could gain confidence in that we were indeed fractioning the search space into an adequate number of communities. We also diagnosed our choice of two communities by reporting the rank (from lowest to largest) of the first eigenvalue produced by SP to be non-zero. This eigenvalue serves as an estimator of the optimal number of communities in a network [@Chung1999], such that a rank of two in our partitions justifies limiting the number of communities to two.

>>>>

``` {r fMRI data descriptives} 
# load file with the n of TRs and density data
corrCounts <- read.csv('corrCounts.csv')
```
Our fMRI analyses used resting-state fMRI data from the Human Connectome Project [@VanEssen2012] Q6 release (n = `r nSubj`, randomly sampled from the total pool of 469 available subjects). Each subject's data was acquired over two days at Washington University in St. Louis on a Siemens CONNECTOM Skyra MRI scanner (Siemens, Erlangen, Germany). Four resting state runs (TR = 0.720 s, TE = 33.1 ms, FA = 52°, multiband factor = 8, 72 slices, 2 mm isotropic voxels) each comprised up to 1200 time points(14 min 33 s) for a total of up to 4800 time points (median = `r median(corrCounts$nTR)`, minimum = `r min(corrCounts$nTR)`). Two runs were acquired on each day, with the phase encoding direction set to left-right for one run and right-left for the other. Only subjects with both left-right and right-left phase encoding were included (i.e. subjects with four rsfMRI sessions). In addition, only datasets with either low motion levels (under 1.5 mm) or under 0.5 mm mean framewise displacement (FD) were used. See [@VanEssen2012] for more details about the data acquisition protocol.

@@

```{r Density calculation}
# Load density data
corrCounts <- corrCounts %>%
  mutate(prcntNS = NonSignificant / Total,
         prcntPos = SignificantPos / (Total - NonSignificant),
         prcntNeg = SignificantNeg / (Total - NonSignificant),
         density = 1 - prcntNS)

# density descriptives
mdAll <- mean(corrCounts$density)
sddAll <- sd(corrCounts$density)

# prcnt pos/neg
mPos <- round(mean(corrCounts$prcntPos), digits = 2)
sdPos <-round(sd(corrCounts$prcntPos), digits = 2)
mNeg <- round(mean(corrCounts$prcntNeg), digits = 2)
sdNeg <-round(sd(corrCounts$prcntNeg), digits = 2)
```

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). The proportion of remaining edges (i.e. network density) was high for all individuals (mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`), and was mostly composed of positive correlations (positive: mean = `r mPos`, sd = `r sdPos`; negative: mean = `r mNeg`, sd = `r sdNeg`). Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. We generated and analyzed network weight matrices at four levels: (1) for each run separately (1200 TRs each); (2) for the concatenated time series for the two runs on each day (2400 TRs total); (3) for each subject's full dataset (4800 TRs total); and (4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

In order to evaluate the convergent validity of the resulting partitionings across community-detection methods, we also estimated these communities using the more traditional approach of modularity maximization, based on the algorithm from Clauset et al. [-@Clauset2004]. This method compares the strength of the connection between any two vertices to their probability of being randomly connected to any other vertex in the network. The method heuristically iterates through many possible combinations of vertices, and selects the partitioning that maximizes the network's so-called modularity. Unlike SP, modularity can fractionate a network into more than two communities. We assumed that if the partitionings from the bounded and unbounded community detection methods showed high levels of agreement, we could gain confidence in that we were indeed fractioning the search space into an adequate number of communities. We also diagnosed our choice of two communities by reporting the rank (from lowest to largest) of the first eigenvalue produced by SP to be non-zero. This eigenvalue serves as an estimator of the optimal number of communities in a network [@Chung1999], such that a rank of two in our partitions justifies limiting the number of communities to two.
